Representation Learning for Generalized Planning - SAT Encodings

### 2020-09-05 doc/encodings/paper.tex
Update PDF encoding

\documentclass[a4paper]{article}
%\def\year{2020}\relax
%\usepackage{aaai20}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS

% our packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{bm}
\usepackage{booktabs}
%\usepackage[inline]{enumitem}
%\usepackage{mathtools}
%\usepackage{multirow}
%\usepackage[mode=buildnew,subpreambles=true]{standalone}
\usepackage{subcaption}
%\usepackage{todonotes}
\usepackage[table]{xcolor}  % TODO Comment out for final submission
%\usepackage{pgfplots}
%\usepackage{tikz}
\usepackage{natbib}

@@

 \item $V(s) \leq d$ for $s$ alive, and $d \in [0, D]$, where $D = \max_{s} \delta \cdot V^*(s)$.
 Note that for states $s$ that are a goal, we know $V(s)=0$, hence $V(s) \leq d$ for all $d$,
 and for states $s$ that are unsolvable, we know that $V(s) \not\leq d$ for all $d$.
 Thus, we can restrict SAT variables $V(s) \leq d$ to those states $s$ that are alive.

@@

\subsubsection{Auxiliary Variables}
\begin{itemize}
 \item We'll use $GV(s, s', d)$ to denote that $Good(s, s')$ and $V(s') \leq d$. This is enforced with the constraint
 $GV(s, s', d) \rightarrow Good(s, s') \land V(s') \leq d$ for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.
 The other direction of the implication is not necessary.
\end{itemize}

@@

\smallpar{C2-3} Labeling a transition $(s, s')$ as ``good'' necessarily upper-bounds $V(s)$:
\begin{align}
 Good(s, s') \land V(s') \leq d \rightarrow V(s) \leq d+1,&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.} \\
 Good(s, s') \rightarrow V(s) \leq d+1,&\;\; \text{for $s$ alive, $s'$ goal, $(s, s') \not\in \badtx$, $d \in [0, D)$.}
\end{align}

\smallpar{C3'} All transitions chosen as \emph{Good} need to have the same q-value:
\begin{align}
 V(s) \leq d+1 \land \neg V(s') \leq d \rightarrow \neg Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.} \tag{\theequation${}^\prime$}
\end{align}

\smallpar{C4-5} Any upper bound on $V(s)$ (for $s$ not a goal) needs to be justified:
\begin{align}
 V(s) \leq d+1 \rightarrow \bigvee_{\substack{
 s' \text{ goal child of } s\\
 (s, s') \not\in \badtx}} Good(s, s') \lor
 \bigvee_{\substack{
 s' \text{ alive child of } s\\
 (s, s') \not\in \badtx}} GV(s, s', d),&
 \;\; \text{for $s$ alive, $d \in [0, D)$.} \\
 \neg V(s) \leq 0,&\;\; \text{for $s$ not a goal.}
\end{align}

\smallpar{C6}
At least one atom $V(s) \leq d$ is true for each alive state $s$:
\begin{align}
 \bigvee_{d \in [0, D]} V(s) \leq d,&\;\; \text{for $s$ alive.}
\end{align}

\smallpar{C7}
Atoms $V(s) \leq d$ are consistent among them:
\begin{align}
 V(s) \leq d \rightarrow V(s) \leq d+1,&\;\; \text{for $s$ alive, $d \in [0, D)$.}
\end{align}

\smallpar{C8-9}
Good transitions can be distinguished from bad transitions.
Let $(s, s')$ and $(t, t')$ be \emph{representative} transitions
of two different equivalence classes such that $(s, s') \not\in \badtx$
(which implies that $s'$ is solvable). Then,

@@

\noindent where $Dist(s, s', t, t')$ is shorthand for $\bigvee_{f \in D2(s, s', t, t')} Select(f)$.

\smallpar{C10}
$V(s)$ cannot be much larger than $V^+(s)$:
\begin{align}
 V(s) \leq \delta \cdot V^*(s),&\;\; \text{for $s$ alive.}
\end{align}

@@

>>>>

\documentclass[a4paper]{article}
%\def\year{2020}\relax
%\usepackage{aaai20}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\frenchspacing  % DO NOT CHANGE THIS
%\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
%\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS

\usepackage[margin=1in]{geometry}

% our packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{bm}
\usepackage{booktabs}
%\usepackage[inline]{enumitem}
%\usepackage{mathtools}
%\usepackage{multirow}
%\usepackage[mode=buildnew,subpreambles=true]{standalone}
\usepackage{subcaption}
%\usepackage{todonotes}
\usepackage[table]{xcolor}  % TODO Comment out for final submission
%\usepackage{pgfplots}
%\usepackage{tikz}
\usepackage{natbib}
\usepackage{fancyvrb}

@@

 \item $V(s, d)$ for $s$ alive, and $d \in [0, D]$, where $D = \max_{s} \delta \cdot V^*(s)$,
 with intended denotation $V(s)=d$.
 Note that for states $s$ that are a goal, we know $V(s)=0$,
 and for states $s$ that are unsolvable, we know that $V(s) \neq d$ for all $d$.
 Thus, we can restrict SAT variables $V(s, d)$ to those states $s$ that are alive.

@@

%\subsubsection{Auxiliary Variables}
%\begin{itemize}
% \item We'll use $GV(s, s', d)$ to denote that $Good(s, s')$ and $V(s') \leq d$. This is enforced with the constraint
% $GV(s, s', d) \rightarrow Good(s, s') \land V(s') \leq d$ for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [0, D)$.
% The other direction of the implication is not necessary.
%\end{itemize}

@@

\smallpar{C2} $V$ is always descending along Good transitions:
\begin{align}
 Good(s, s') \land V(s', d) \rightarrow \bigvee_{d < k \leq D} V(s, k),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $d \in [1, D)$.} \\
 V(s', D) \rightarrow \neg Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$.} \tag{\theequation${}^\prime$}
\end{align}

%\smallpar{C4-5} Any upper bound on $V(s)$ (for $s$ not a goal) needs to be justified:
%\begin{align}
% V(s) \leq d+1 \rightarrow \bigvee_{\substack{
% s' \text{ goal child of } s\\
% (s, s') \not\in \badtx}} Good(s, s') \lor
% \bigvee_{\substack{
% s' \text{ alive child of } s\\
% (s, s') \not\in \badtx}} GV(s, s', d),&
% \;\; \text{for $s$ alive, $d \in [0, D)$.} \\
% \neg V(s) \leq 0,&\;\; \text{for $s$ not a goal.}
%\end{align}

\smallpar{C3-4}
Variables $V(s, d)$ define a function that is total over the set of alive states,
and such that $V(s)$ is within lower bound $V^*(s)$ and upper bound $\delta \cdot V^*(s)$:
\begin{align}
 \bigvee_{V^*(s) \leq d \leq \delta \cdot V^*(s)} V(s,d),&\;\; \text{for $s$ alive.} \\
 \neg V(s, d) \lor \neg V(s, d')&\;\; \text{for $s$ alive, $1 \leq d < d' \leq D$.}
\end{align}

\smallpar{C5-6}
Good transitions can be distinguished from bad transitions.
Let $(s, s')$ and $(t, t')$ be \emph{representative} transitions
of two different equivalence classes such that $(s, s') \not\in \badtx$
(which implies that $s'$ is solvable). Then,

@@

\noindent where $Dist(s, s', t, t')$ is shorthand for $\bigvee_{f \in D1\&2(s, s', t, t')} Select(f)$.

\smallpar{C7 (Optional)}
Goals are distinguishable from non-goals.
\begin{align}
\bigvee_{f \in D1(s, s')} Select(f),&\;\; \text{for $s$ goal, $s$ not a goal}
\end{align}

\smallpar{C8 (Optional)}
All selected features need to have some Good transition that takes them to $0$:
\begin{align}
 Selected(f) \rightarrow \bigvee_{(s, s') \in Z(f)} Good(s, s'),&\;\; \text{for $f$ in pool}
\end{align}

\noindent where $Z(f)$ is the set transitions starting in an alive state that change the denotation of $f$ from something
larger than 0 to 0.

@@

\section{Empirical Results}

\subsection{Blocks:clear}
\begin{itemize}
 \item Two instances of 5 blocks.
 \item 1 iteration, 19759 vars + 69255 clauses
 \item Number of transitions/equivalence classes: 1161/55
 \item Optimal MAXSAT solution with cost 5 found
\end{itemize}

\begin{Verbatim}[fontsize=\footnotesize]
0. n(a) DECs AND n-ontable NILs AND n-ontable>0
1. n(a) NILs AND n-ontable INCs AND n-ontable>0
\end{Verbatim}

\subsection{Blocks:on}
\begin{Verbatim}[fontsize=\footnotesize]
 1. clear(a) NILs AND clear(b) ADDs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
 2. clear(a) ADDs AND clear(b) NILs AND clear(b)=0 AND holding(·) DELs AND n-clear INCs AND n-clear>0
 3. clear(a) ADDs AND clear(b) DELs AND holding(·) DELs AND n-clear NILs AND n-clear>0
 4. clear(a) DELs AND clear(b) NILs AND clear(b)>0 AND holding(·) ADDs AND n-clear DECs
 5. clear(a) NILs AND clear(a)=0 AND clear(b) DELs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
 6. clear(a) ADDs AND clear(b) NILs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
 7. clear(a) NILs AND clear(b) NILs AND holding(·) DELs AND n-clear INCs AND n-clear>0
 8. clear(a) ADDs AND clear(b) DELs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
 9. clear(a) DELs AND clear(b) NILs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
 10. clear(a) NILs AND clear(b) ADDs AND holding(·) DELs AND n-clear INCs AND n-clear>0
 11. clear(a) NILs AND clear(a)=0 AND clear(b) NILs AND holding(·) ADDs AND n-clear NILs AND n-clear>0
 12. clear(a) NILs AND clear(a)>0 AND clear(b) NILs AND clear(b)=0 AND holding(·) ADDs AND n-clear NILs AND n-clear>0
\end{Verbatim}

\subsection{Blocks:arbitrary}
(Atomic move encoding, experiment name: all$\_$at$\_$5)

\begin{itemize}
 \item 1 instance 5 blocks
 \item 11 iterations
 \item CNF Theory: 21795 vars + 189561 clauses
 \item 12659 features
 \item concept size bound=10,
 \item Number of transitions/equivalence classes: 11053/133
 \item Optimal MAXSAT solution with cost 9 found
\end{itemize}

\begin{Verbatim}[fontsize=\footnotesize]
1. n-clear INCs AND n-clear>0 AND n-ontarget DECs AND n-superficially-well-placed NILs AND n-superficially-well-placed>0
2. n-clear NILs AND n-clear>0 AND n-ontarget INCs AND n-ontarget>0 AND n-superficially-well-placed INCs AND n-superficially-well-placed>0
3. n-clear INCs AND n-clear>0 AND n-ontarget NILs AND n-ontarget>0 AND n-superficially-well-placed NILs AND n-superficially-well-placed>0
4. n-clear NILs AND n-clear>0 AND n-ontarget NILs AND n-ontarget>0 AND n-superficially-well-placed INCs AND n-superficially-well-placed>0
5. n-clear INCs AND n-clear>0 AND n-ontarget INCs AND n-ontarget>0 AND n-superficially-well-placed INCs AND n-superficially-well-placed>0
6. n-clear DECs AND n-ontarget INCs AND n-ontarget>0 AND n-superficially-well-placed INCs AND n-superficially-well-placed>0
7. n-clear INCs AND n-clear>0 AND n-ontarget NILs AND n-ontarget>0 AND n-superficially-well-placed DECs
\end{Verbatim}

\subsection{Gripper}
One simple (smallest) standard instance, 4 blocks, two rooms.
1 iteration
CNF Theory: 34418 vars + 124935 clauses
835 features
Number of transitions/equivalence classes: 1140/65
concept size bound=10,
Optimal MAXSAT solution with cost 5 found

\begin{Verbatim}[fontsize=\footnotesize]
0. ncarried NILs AND ncarried=0 AND robby-at-B DECs
1. ncarried DECs AND robby-at-B NILs AND robby-at-B>0
2. ncarried INCs AND ncarried=0 AND robby-at-B NILs AND robby-at-B=0
3. ncarried NILs AND ncarried>0 AND robby-at-B INCs AND robby-at-B=0
\end{Verbatim}

\subsection{Spanner}

\begin{itemize}
 \item 8 instances with different spanner configurations, goal is always the same.
 \item 1 iteration
 \item CNF Theory: 466575 vars + 1732026 clauses
 \item 724 features
 \item concept size bound=8,
 \item Number of transitions/equivalence classes: 11053/133
 \item Optimal MAXSAT solution with cost 9 found
\end{itemize}

NOTE: Features of the form $f1>f2$ are not necessary for this solution.

\begin{Verbatim}[fontsize=\footnotesize]
0. bobs-loc-empty NILs AND bobs-loc-empty=0 AND n-items-somewhere DECs AND n-tightened NILs AND n-tightened=0
1. bobs-loc-empty NILs AND bobs-loc-empty>0 AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened NILs AND n-tightened=0
2. bobs-loc-empty DELs AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened NILs AND n-tightened=0
3. bobs-loc-empty NILs AND bobs-loc-empty=0 AND n-items-somewhere NILs AND n-items-somewhere>0 AND n-tightened INCs
4. bobs-loc-empty ADDs AND n-items-somewhere DECs AND n-tightened NILs AND n-tightened=0
\end{Verbatim}

\subsection{Pick Reward}
1 5x5 training instance
3 iteration
CNF Theory: 18519 vars + 76267 clauses
514 features
Number of transitions/equivalence classes: 432/361
concept size bound=8,
Optimal MAXSAT solution with cost 7 found

\begin{Verbatim}[fontsize=\footnotesize]
 1. n-rewards NILs AND n-rewards>0 AND unblocked-dist-to-closest-reward DECs AND unblocked-dist-to-closest-reward>0
 2. n-rewards DECs AND unblocked-dist-to-closest-reward INCs AND unblocked-dist-to-closest-reward=0
\end{Verbatim}
